---
title: "Colour Palettes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Colour Palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 3, dpi = 120
)
```

```{r}
library(UtilsR)
library(ggplot2)
```

## Overview

UtilsR provides a palette system with 256+ colour schemes from
RColorBrewer, ggsci, viridis, rcartocolor, nord, and more.

**Core components:**

| Function | Purpose |
|----------|---------|
| `pal_lancet` | Default palette object (15 Lancet journal colours) |
| `palette_list` | Named list of 256 palettes |
| `pal_get()` | Extract colours from any palette by name |
| `pal_show()` | Visualise a palette as colour swatches |
| `pal_list()` | Browse all available palette names |
| `show_color()` | Display any colour vector in console |
| `as_palette()` | Create custom palette objects |

---

## `pal_lancet` — Default Palette

The built-in default palette used by `plt_cat()` and other plot functions:

```{r lancet}
as.character(pal_lancet)
as.character(pal_lancet[1:5])
```

```{r lancet-plot}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_manual(values = as.character(pal_lancet[1:3])) +
  theme_my()
```

---

## `palette_list` — 256 Palettes

All palettes are stored in `palette_list` (internal data, accessed via `pal_get()`):

```{r palette-list}
# Browse available palette names
head(names(UtilsR:::palette_list), 20)
length(UtilsR:::palette_list)
```

---

## `pal_get()` — Extract Palette Colours

Retrieve colours from any palette by name:

```{r pal-get}
# Get 5 colours from "Paired"
pal_get("Paired", n = 5)

# Get 3 colours from "Set1"
pal_get("Set1", n = 3)

# Get 8 colours from "Dark2"
pal_get("Dark2", n = 8)
```

Use in ggplot2:

```{r pal-get-plot}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_manual(values = as.character(pal_get("Set2", n = 3))) +
  theme_my()
```

---

## `pal_show()` — Visualise Palettes

Display colour swatches for any palette:

```{r pal-show, eval=FALSE}
pal_show("Paired")
pal_show("Set1")
```

---

## `pal_list()` — Browse All Palettes

```{r pal-list, eval=FALSE}
# Interactive palette browser (opens viewer)
pal_list()
```

---

## `show_color()` — Display Any Colours

```{r show-color, eval=FALSE}
show_color(c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FFEAA7"))
```

---

## `as_palette()` — Custom Palettes

Create your own palette object with auto-display:

```{r as-palette}
my_pal <- as_palette(c("#264653", "#2A9D8F", "#E9C46A", "#F4A261", "#E76F51"))
as.character(my_pal)  # display as plain hex vector
```

---

## Use with `plt_cat()`

Palettes integrate seamlessly with plot functions:

```{r plt-cat-named, fig.height=4}
set.seed(1)
df <- data.frame(Type = factor(sample(LETTERS[1:5], 200, TRUE)))

# Use named palette
plt_cat(df, "Type", type = "pie", label = TRUE, palette = "Set2")
```

```{r plt-cat-custom, fig.height=4}
# Use custom colours
plt_cat(df, "Type", type = "bar", stat = "count",
        palette = c("#264653", "#2A9D8F", "#E9C46A", "#F4A261", "#E76F51"))
```

```{r plt-cat-lancet, fig.height=4}
# Default (pal_lancet) — no palette argument needed
plt_cat(df, "Type", type = "pie", label = TRUE)
```
