---
title: "Colour Palettes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Colour Palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 3, dpi = 120
)
```

```{r}
library(UtilsR)
library(ggplot2)
```

## Built-in Palettes

UtilsR provides 11 colour palettes (16--48 colours each):

| Palette | N | Source |
|---------|---|--------|
| `pal_lancet` | 15 | Lancet journal |
| `pal_ditto` | 40 | dittoSeq (scRNA-seq) |
| `pal_igv` | 48 | IGV genome browser |
| `pal_polychrome` | 36 | Max perceptual distinctness |
| `pal_glasbey` | 32 | Glasbey algorithm |
| `pal_alphabet` | 26 | Green-Armytage A--Z |
| `pal_ucsc` | 26 | UCSC genome browser |
| `pal_kelly` | 22 | Kelly max contrast |
| `pal_d3` | 20 | D3.js Category20 |
| `pal_trubetskoy` | 20 | Trubetskoy distinct |
| `pal_simpsons` | 16 | Simpsons ggsci |

## Usage

### Direct access

```{r direct}
pal_lancet[1:5]
```

### In ggplot2

```{r ggplot}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  scale_color_manual(values = pal_lancet[1:3]) +
  theme_my()
```

### Browse all palettes

```{r pal-list, eval=FALSE}
pal_list()
```

### Get specific palette by name

```{r pal-get, eval=FALSE}
pal_get("Paired", n = 6)
pal_get("Set1", n = 3)
```

### Create custom palette

```{r as-palette}
my_pal <- as_palette(c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4"))
my_pal
```

### Use in plt_cat

```{r plt-cat-palette, fig.height=4}
set.seed(1)
df <- data.frame(Type = factor(sample(LETTERS[1:4], 200, TRUE)))
plt_cat(df, "Type", type = "pie", label = TRUE,
        palette = c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4"))
```
