---
title: "ggplot2 Formatting: fmt_* and Themes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggplot2 Formatting: fmt_* and Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120,
  message = FALSE, warning = FALSE
)
```

```{r}
library(UtilsR)
library(ggplot2)
```

## Base Plot

All examples build on this base plot:

```{r base}
p <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point(size = 2)
p
```

---

## `fmt_plot()` — Master Chaining

Chain multiple formatting operations in one call:

```{r fmt-plot}
p |> fmt_plot(
  legend.position = "bottom",
  tag = "A",
  base_size = 14
)
```

---

## `fmt_tag()` — Panel Labels

```{r fmt-tag}
p |> fmt_tag("A")
p |> fmt_tag("B", x = 0.95, y = 0.95, size = 16)
```

---

## `fmt_legend()` — Legend Formatting

```{r fmt-legend}
p |> fmt_legend(position = "bottom", direction = "horizontal")
```

---

## `fmt_ref()` — Reference Lines

```{r fmt-ref}
p |> fmt_ref(xintercept = 5.8, yintercept = 3.0)
```

```{r fmt-ref-multi}
# Multiple lines with colours
p |> fmt_ref(xintercept = c(5, 6, 7), color = c("red", "blue", "green"))
```

---

## `fmt_axis()` — Axis Control

```{r fmt-axis}
# Hide x-axis (useful for multi-panel layouts)
p |> fmt_axis(x = FALSE)
```

---

## `fmt_strip()` — Facet Strip Colours

```{r fmt-strip, fig.width=8}
p_facet <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point() +
  facet_wrap(~Species)

p_facet |> fmt_strip(colors = c("#E41A1C", "#377EB8", "#4DAF4A"))
```

---

## `fmt_bg()` — Background Stripes

```{r fmt-bg}
p |> fmt_bg(colors = c("grey95", "white"), n = 3, direction = "x")
```

---

## `fmt_scale()` — Axis Scales

```{r fmt-scale}
p |> fmt_scale(x_limits = c(4, 8), y_limits = c(2, 4.5))
```

---

## `fmt_boxplot()` — Overlay Boxplot

```{r fmt-boxplot}
p_violin <- ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_violin()
p_violin |> fmt_boxplot()
```

---

## Themes

### `theme_my()` — Clean General Purpose

```{r theme-my}
p + theme_my()
```

### `theme_my()` with custom base size

```{r theme-my-large}
p + theme_my(base_size = 16)
```

---

## `flatten_patchwork()` — Flatten Nested Patchwork

```{r flatten, eval=FALSE}
library(patchwork)
nested <- (p1 | p2) / (p3 | p4)
flat <- flatten_patchwork(nested)
```

Recursively flattens nested patchwork objects into a single-level list.
