---
title: "plt_cat: Unified Categorical Plot Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plt_cat: Unified Categorical Plot Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 8, fig.height = 5, dpi = 120,
  message = FALSE, warning = FALSE
)
```

`plt_cat()` is a single entry point for **11 categorical chart types** with 27+ parameters.
It handles grouping, splitting, dodge/stack, labels, background bands, NA handling, and more.

## Sample Data

```{r data}
library(UtilsR)
library(ggplot2)

set.seed(1)
df <- data.frame(
  Type  = factor(sample(c("A","B","C"), 200, TRUE, c(0.5, 0.3, 0.2))),
  Group = factor(sample(c("X","Y","Z"), 200, TRUE)),
  Batch = factor(sample(c("B1","B2"), 200, TRUE))
)
df$Region <- factor(ifelse(df$Group %in% c("X","Y"), "R1", "R2"))
head(df)
```

---

## 1. Bar Chart

The most common type. Supports `stat = "percent"` (default) or `"count"`,
and `position = "stack"` (default) or `"dodge"`.

```{r bar-stack}
plt_cat(df, "Type", "Group", type = "bar")
```

```{r bar-dodge}
plt_cat(df, "Type", "Group", type = "bar",
        stat = "count", position = "dodge", label = TRUE)
```

```{r bar-flip}
plt_cat(df, "Type", "Group", type = "bar",
        flip = TRUE, label = TRUE)
```

### No group (single variable)

When `group.by` is omitted, each stat.by level becomes its own bar:

```{r bar-nogroup}
plt_cat(df, "Type", type = "bar", stat = "count",
        label = TRUE, legend.position = "none")
```

### Background bands

Use `bg.by` in dodge mode to add alternating background colour bands.
`bg.by` must be a superset of `group.by` (each group belongs to exactly one bg level).

```{r bar-bg}
plt_cat(df, "Type", "Group", type = "bar",
        position = "dodge", bg.by = "Region")
```

```{r bar-bg-custom}
plt_cat(df, "Type", "Group", type = "bar",
        position = "dodge", bg.by = "Region",
        bg_palette = c("lightblue", "lightyellow"),
        title = "Custom Background Colors")
```

---

## 2. Rose Chart

A bar chart in polar coordinates (`theta = "x"`). Good for cyclical data.

```{r rose}
plt_cat(df, "Type", "Group", type = "rose")
```

```{r rose-dodge}
plt_cat(df, "Type", "Group", type = "rose",
        stat = "count", position = "dodge")
```

---

## 3. Ring (Donut) Chart

A pie chart with a hole in the center. Supports dodge for grouped spiral display.

```{r ring}
plt_cat(df, "Type", type = "ring", label = TRUE)
```

```{r ring-dodge}
plt_cat(df, "Type", "Group", type = "ring",
        stat = "count", position = "dodge")
```

---

## 4. Pie Chart

```{r pie}
plt_cat(df, "Type", type = "pie", label = TRUE)
```

```{r pie-styled}
plt_cat(df, "Type", type = "pie",
        palette = "Paired",
        title = "Type Distribution",
        subtitle = "n = 200",
        legend.position = "bottom",
        legend.direction = "horizontal")
```

---

## 5. Trend Chart

Combines stepped area background with bar overlay. Best for showing trends across ordered groups.

```{r trend}
plt_cat(df, "Type", "Group", type = "trend", stat = "count")
```

```{r trend-label}
plt_cat(df, "Type", "Group", type = "trend",
        stat = "count", label = TRUE)
```

---

## 6. Area Chart

Stacked area chart. Uses continuous x-axis internally.

```{r area}
plt_cat(df, "Type", "Group", type = "area", stat = "count")
```

```{r area-dodge}
plt_cat(df, "Type", "Group", type = "area",
        stat = "count", position = "dodge")
```

---

## 7. Dot Plot

Bubble-style dot plot where size encodes value.

```{r dot}
plt_cat(df, "Type", "Group", type = "dot")
```

```{r dot-label}
plt_cat(df, "Type", "Group", type = "dot",
        label = TRUE, stat = "count")
```

---

## 8. Sankey Diagram

Requires `stat.by` with >= 2 column names. Shows flow between categories.

```{r sankey}
plt_cat(df, c("Type", "Group"), type = "sankey")
```

```{r sankey-3}
plt_cat(df, c("Type", "Group", "Batch"), type = "sankey")
```

---

## 9. Chord Diagram

Shows pairwise connections between two categorical variables.
Returns base R graphics (not ggplot).

```{r chord, fig.height=6}
plt_cat(df, c("Type", "Group"), type = "chord")
```

---

## 10. Venn Diagram

Shows set intersections. Use `stat_level` to specify which level counts as "positive".

```{r venn-binary}
set.seed(42)
df_bin <- data.frame(
  A = sample(c("Yes","No"), 200, TRUE),
  B = sample(c("Yes","No"), 200, TRUE),
  C = sample(c("Yes","No"), 200, TRUE)
)
plt_cat(df_bin, c("A","B","C"), type = "venn")
```

```{r venn-stat-level}
plt_cat(df, c("Type", "Group"), type = "venn",
        stat_level = c("A", "X"))
```

---

## 11. UpSet Plot

A scalable alternative to Venn diagrams for showing set intersections.

```{r upset-binary}
plt_cat(df_bin, c("A","B","C"), type = "upset")
```

```{r upset-stat-level}
plt_cat(df, c("Type", "Group"), type = "upset",
        stat_level = c("A", "X"))
```

---

## Cross-cutting Features

### split.by

Split data by a variable and create one panel per level:

```{r split-bar}
plt_cat(df, "Type", "Group", split.by = "Batch", type = "bar")
```

```{r split-pie}
plt_cat(df, "Type", split.by = "Batch", type = "pie",
        facet_ncol = 2)
```

### NA Handling

Include NA values in the plot with `NA_stat = TRUE`:

```{r na-stat}
df_na <- df
df_na$Type[1:20] <- NA
plt_cat(df_na, "Type", "Group", type = "bar",
        NA_stat = TRUE, NA_color = "red")
```

### keep_empty

Preserve dropped factor levels as empty bars:

```{r keep-empty}
df2 <- df[df$Type != "C", , drop = FALSE]
plt_cat(df2, "Type", "Group", type = "bar", keep_empty = TRUE)
```

### Label Styling

Customise label appearance with `label.fg`, `label.bg`, and `label.bg.r`:

```{r label-style}
plt_cat(df, "Type", "Group", type = "bar",
        stat = "count", position = "dodge",
        label = TRUE, label.fg = "white", label.bg = "black")
```

### aspect.ratio

Control panel aspect ratio (auto-set to 1 for polar types):

```{r aspect}
plt_cat(df, "Type", "Group", type = "bar", aspect.ratio = 0.5)
```

### Custom Palette

```{r palette}
plt_cat(df, "Type", type = "pie",
        palette = c("#FF6B6B", "#4ECDC4", "#45B7D1"),
        label = TRUE)
```

---

## Parameter Reference

| Group | Parameters |
|-------|-----------|
| **Data** | `stat.by`, `group.by`, `split.by` |
| **Chart** | `type` (11 types), `stat`, `position` |
| **Colour** | `palette`, `alpha`, `NA_color` |
| **Labels** | `label`, `label.size`, `label.fg`, `label.bg`, `label.bg.r` |
| **Background** | `bg.by`, `bg_palette`, `bg_alpha` |
| **NA/Empty** | `NA_stat`, `keep_empty` |
| **Layout** | `title`, `subtitle`, `xlab`, `ylab` |
| **Legend** | `legend.position`, `legend.direction` |
| **Aspect** | `aspect.ratio` |
| **Split** | `facet_nrow`, `facet_ncol`, `facet_byrow` |
| **Set types** | `stat_level` |
| **Safety** | `force` |
